---
title: "An Introduction to fingertipsR and fingertipscharts"
author: "Sebastian Fox\nPublic Health Data Science, PHE"
output:
  phecharts::html_phe_slidy
---

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(fingertipsR)
opts_chunk$set(echo=FALSE,
               cache=FALSE)
```

## Outline

- [Fingertips](https://fingertips.phe.org.uk/) overview
- Interactive workshop part 1 - `fingertipsR`
- Break
- Interactive workshop part 2 - `fingertipscharts`

## Fingertips
[Fingertips](https://fingertips.phe.org.uk/) is a platform containing 1,700 indicators of public health in England.

It presents data at different geographies, and in many cases can be cut in different age groups, sexes and time period.

For many indicators, significance testing is performed to give added insight into how areas are performing when compared to their peers.

It presents the data in many different visualisations that can be used for different purposes.

<iframe src="https://fingertips.phe.org.uk/" width='90%' height='720px'></iframe>


## fingertipsR

### What does the package do?

Simply put, the package translates user inputs into a url, which maps to the Fingertips API. It then imports data from the Application Programming Interface (API) into the users R workspace.

The package accesses the internet, so a network connection is required when you use it. Also, the amount of data you are importing will affect the time it takes to import it.

This is a typical url generated:

https://fingertips.phe.org.uk/api/all_data/csv/by_indicator_id?indicator_ids=10101&child_area_type_id=202&parent_area_type_id=6&include_sortable_time_periods=yes

This is what the data at that url looks like:

<iframe src="https://fingertips.phe.org.uk/api/all_data/csv/by_indicator_id?indicator_ids=10101&child_area_type_id=202&parent_area_type_id=6&include_sortable_time_periods=yes" width='90%' height='720px'></iframe>


## The data provided from fingertipsR

This table is what the previous data comes into your R environment looking like:

```{r import_data}
df <- fingertips_data(IndicatorID = 10101,
                      AreaTypeID = 202)

DT::datatable(df)
```

## fingertips_data: the main function

These are the arguments that the user can provide to `fingertips_data()`

```{r fingertips_data, eval=FALSE, echo=TRUE}
fingertips_data(IndicatorID = NULL, 
                AreaCode = NULL, 
                DomainID = NULL,
                ProfileID = NULL, 
                AreaTypeID, 
                ParentAreaTypeID = NULL, 
                categorytype = FALSE,
                rank = FALSE, 
                url_only = FALSE, 
                path) 
```

## Common problems

- `AreaTypeID` mismatch with `IndicatorID`
- Multiple age groups/sex within data, therefore analysis not adding up to what is expected
- Script worked one day but not the week after; Fingertips updates its data on the 1st Tuesday of each month. Sometimes this means indicators are removed or changed. The API is live...
- What does `ParentAreaTypeID` refer to? See "Areas grouped by" on Fingertips website
- Any other issues, please check here and report them if they aren't covered: https://github.com/ropensci/fingertipsR/issues

## Accessing the documentation

To access the documentation for a function, put a question mark in front of it within your console: `?fingertips_data`

(hint: you can also double click on the function in a script and press F1 on the keyboard)



